Confirmation of \srcfull{} as a long period radio transient

### 2025-05-18 paper/main.tex
Wrote summary in paper. Synched with Overleaf

In this short paper, we present new radio observations (spanning frequencies from $154\,$MHz to $1.7\,$GHz) in which \src{} was re-detected, confirming that it is indeed an LPT with a period of ${\sim}1.16\,$hours.
In \S\ref{sec:observations}, we describe the observations themselves, some of which were taken as part of independent surveys, and some of which were taken as dedicated follow-up of \src{} once it became apparent that it was still active.
In \S\ref{sec:analysis} we present a timing analysis of the new detections, as well as revised estimates of the DM and the scattering timescale.
We also include new polarization detections in which the polarisation angle (PA) behavior differs significantly from the originally detected pulse.
In \S\ref{sec:discussion} we revisit the question of the nature of \src{}, arguing in favor of a white dwarf interpretation.
The main conclusions are summarized in \S\ref{sec:summary}.

@@

The measured DM of $764 \pm 25\,{\rm pc}\,{\rm cm}^{-3}$ is slightly higher than, but still consistent with, the in-band measurement of $710^{+200}_{-180}\,{\rm pc}\,{\rm cm}^{-3}$ given in \citetalias{2024MNRAS.535..909D}.
Because we have not attempted to simultaneously fit for both DM and scattering timescale, there will remain some small systematic error on this reported DM measurement.
Breaking the degeneracy is possible in principle with a much higher S/N average profile than what is achievable with the current data set.

For planning observations of \src{} at low ($\lesssim 300 MHz$), care should be taken to incorporate scattering into the predicted ToAs.
The $N$th pulse since PEPOCH is predicted to arrive at the Solar System barycentre at
\begin{equation}
    {\rm MJD} = {\rm PEPOCH} + NP + \Delta t_{\rm DM}(\nu) + \Delta t_{\rm sc}(\nu),
\end{equation}
where PEPOCH and the period, $P$, are given in \Tab~\ref{tbl:ephemeris},
\begin{equation}
    \Delta t_{\rm DM}(\nu) = 4148\,{\rm s} \, \left(\frac{\rm DM}{{\rm pc\,cm}^{-3}}\right) \left(\frac{\nu}{\rm MHz}\right)^{-2}
\end{equation}
is the usual dispersion attributed to the interstellar medium, and
\begin{equation}
    \Delta t_{\rm sc}(\nu) = -\sqrt{2}\sigma\erfcx^{-1}\left(\frac{\tau(\nu)}{\sigma}\sqrt{\frac{2}{\pi}}\right) + \frac{\sigma^2}{\tau(\nu)}
    \label{eqn:scattering_delay}
\end{equation}
is the amount that the \emph{peak} of the observed pulse is delayed by scattering.
When calculating this quantity, we recommend using $\sigma = 15\,$s, a typical value for low-frequency pulses (see \Fig~\ref{fig:pulse_details}).
At frequencies above $300\,$MHz, $\tau(\nu) \ll \sigma$, and the delay due to scattering, which asymptotically behaves like $\Delta t_{\rm sc}(\nu) \sim \tau$, can be safely neglected.
At frequencies below $150\,$MHz, the asymptotic behaviour
\begin{equation}
    \Delta t_{\rm sc}(\nu) \sim \sigma \sqrt{\ln \left(\frac{\tau^2}{2\pi \sigma^2}\right)} + \frac{\sigma^2}{\tau}
\end{equation}
is accurate to within a few seconds.
At intermediate frequencies ($150\,{\rm MHz} \lesssim \nu \lesssim 300\,{\rm MHz}$), Eq. \eqref{eqn:scattering_delay} should be used directly.

@@

After accounting for dispersion and scattering effects, the timing analysis of \src{} confirms its status as an LPT, as originally conjectured in \citetalias{2024MNRAS.535..909D}.
It has the fifth largest period (${\sim}1.16\,$hours) of known published LPTs, after ASKAP J1839$-$0756 \citep[$6.45\,$hours;][]{Lee2025}, GLEAM-X J0704$-$37 \citep[$2.92\,$hours;][]{2024ApJ...976L..21H,2025A&A...695L...8R}, ILT J1101+5521 \citep[$2.09\,$hours;][]{deRuiter2025}, and GCRT J1745$-$3009 \citep[$1.28\,$hours;][]{2005Natur.434...50H}.

@@

As noted in \citetalias{2024MNRAS.535..909D}, optical follow-up of \src{} is hampered by its proximity to the Galactic plane ($b = -0.12^\circ$).
Inferring this nature of the system (white dwarf vs neutron star, isolated vs binary) from the radio pulses alone is challenging, unless either the population of LPTs is better understood as a whole, or if some other property of the radio signal (e.g. intermittency, polarisation) can be linked to the system properties.

\src{}'s period and duty cycle are not dissimilar to ILT J1101+5521 and GLEAM-X J0704$-$37, which are confirmed polars (\Fig~\ref{fig:lpt_comparison}), suggesting the possibility that \src{}, too, is a WD-MD polar system.
However, its period of ${\sim}1.16\,$hours would make it shorter than the minimum known orbital period of $1.3\,$hours for confirmed polars \citep{schwope2025polarcatcatalogpolarslowaccretion}.
This makes it an interesting testbed for the suggestion advanced by \citet{2025A&A...695L...8R} that ``short LPTs'' ($P \lesssim 78\,$mins) and ``long LPTs'' ($P \gtrsim 78\,$mins) represent different classes of systems (AM CVns and CVs, respectively).
\begin{figure*}{th}
    \centering
    \includegraphics[width=0.98\linewidth]{lpt_comparison.pdf}
    \caption{Duty cycle vs radio pulse period for various classes of radio emitters \citep{deRuiter2025,2022Natur.601..526H,2023Natur.619..487H,2024NatAs...8.1159C,2005Natur.434...50H,deRuiter2025,Lee2025}. The duty cycles are derived from the reported pulse widths for LPTs, which are compared to the range of the fitted, unscattered pulse widths of \src{} at 10\% of the peak. AR\,Sco \citep{2016Natur.537..374M,2018A&A...611A..66S} and other sources without well-defined periods or duty cycles \citep[e.g.][and references therein]{2021ApJ...920...45W} are not included on this plot.}
    \label{fig:lpt_comparison}
\end{figure*}

@@

Another possibility is that radio emission is only observed when the MD produces dense enough particle outflows, in the form of winds, to activate the emission mechanism in the vicinity of the WD.
Such outflows can last for months [CITATION??], and would not be strictly periodic.
If intermittency is of this sort, then the observed activity windows will be similarly aperiodic, or quasiperiodic.
Again, long-term monitoring can help distinguish between these scenarios.

@@

We have shown that \src{} is in fact an LPT, as conjectured in \citetalias{2024MNRAS.535..909D}).
Its period of ${\sim}1.16\,$hours is slightly shorter than the minimum known orbital\dots

>>>>

In this paper, we present new radio observations (spanning frequencies from $154\,$MHz to $1.7\,$GHz) in which \src{} was re-detected, confirming that it is indeed an LPT with a period of ${\sim}1.16\,$hours.
In \S\ref{sec:observations}, we describe the observations themselves, some of which were taken as part of independent surveys, and some of which were taken as dedicated follow-up of \src{} once it became apparent that it was still active.
In \S\ref{sec:analysis} we present a timing analysis of the new detections, as well as revised estimates of the DM and the scattering timescale.
We also include new polarization detections in which the polarisation angle (PA) behavior differs significantly from the originally detected pulse.
In \S\ref{sec:discussion} we revisit the question of the nature of \src{}, arguing in favor of a white dwarf interpretation.
The main conclusions are summarized in \S\ref{sec:summary}.

@@

The measured DM of $764 \pm 25\,{\rm pc}\,{\rm cm}^{-3}$ is slightly higher than, but still consistent with, the in-band measurement of $710^{+200}_{-180}\,{\rm pc}\,{\rm cm}^{-3}$ given in \citetalias{2024MNRAS.535..909D}.
Because we have not attempted to simultaneously fit for both DM and scattering timescale, there will remain some small systematic error on this reported DM measurement.
Note, in particular, that the YMW16 and NE2001 models predict a slightly longer scattering time scale than the $70\,$ms that we have assumed throughout this work.
Breaking the degeneracy is possible in principle with a higher S/N average profile than what is achievable with the current data set.

For planning observations of \src{} at low ($\lesssim 300 MHz$), care should be taken to incorporate scattering into the predicted ToAs.
The $N$th pulse since PEPOCH is predicted to arrive at the Solar System barycentre at
\begin{equation}
    {\rm MJD} = {\rm PEPOCH} + NP + \Delta t_{\rm DM}(\nu) + \Delta t_{\rm sc}(\nu),
    \label{eqn:timing}
\end{equation}
where PEPOCH and the period, $P$, are given in \Tab~\ref{tbl:ephemeris},
\begin{equation}
    \Delta t_{\rm DM}(\nu) = 4148\,{\rm s} \, \left(\frac{\rm DM}{{\rm pc\,cm}^{-3}}\right) \left(\frac{\nu}{\rm MHz}\right)^{-2}
\end{equation}
is the usual dispersion attributed to the interstellar medium, and
\begin{equation}
    \Delta t_{\rm sc}(\nu) = -\sqrt{2}\sigma\erfcx^{-1}\left(\frac{\tau(\nu)}{\sigma}\sqrt{\frac{2}{\pi}}\right) + \frac{\sigma^2}{\tau(\nu)}
    \label{eqn:scattering_delay}
\end{equation}
is the amount that the \emph{peak} of the observed pulse is delayed by scattering.
When calculating this quantity, we recommend using $\sigma = 15\,$s, a typical value for low-frequency pulses (see \Fig~\ref{fig:pulse_details}).
At frequencies above $300\,$MHz, $\tau(\nu) \ll \sigma$, and the delay due to scattering, which asymptotically behaves like $\Delta t_{\rm sc}(\nu) \sim \tau$, can be safely neglected.
At frequencies below $150\,$MHz, the asymptotic behaviour
\begin{equation}
    \Delta t_{\rm sc}(\nu) \sim \sigma \sqrt{\ln \left(\frac{\tau^2}{2\pi \sigma^2}\right)} + \frac{\sigma^2}{\tau}
\end{equation}
is accurate to within a few seconds.
At intermediate frequencies ($150\,{\rm MHz} \lesssim \nu \lesssim 300\,{\rm MHz}$), Eq. \eqref{eqn:scattering_delay} should be used directly.

@@

After accounting for dispersion and scattering effects, the timing analysis of \src{} confirms its status as an LPT, as originally conjectured in \citetalias{2024MNRAS.535..909D}.
It has the fifth largest period (${\sim}1.16\,$hours) of published LPTs to date, after ASKAP J1839$-$0756 \citep[$6.45\,$hours;][]{Lee2025}, GLEAM-X J0704$-$37 \citep[$2.92\,$hours;][]{2024ApJ...976L..21H,2025A&A...695L...8R}, ILT J1101+5521 \citep[$2.09\,$hours;][]{deRuiter2025}, and GCRT J1745$-$3009 \citep[$1.28\,$hours;][]{2005Natur.434...50H}.

@@

As noted in \citetalias{2024MNRAS.535..909D}, optical follow-up of \src{} is hampered by its proximity to the Galactic plane ($b = -0.12^\circ$) and the associated extinction.
Inferring this nature of the system (white dwarf vs neutron star, isolated vs binary) from the radio pulses alone is challenging, until either the population of LPTs is better understood as a whole, or some other property of the radio signal (e.g. intermittency, polarisation) can be linked to the system properties.

\src{}'s period and duty cycle are not dissimilar to ILT J1101+5521 and GLEAM-X J0704$-$37, which are confirmed polars (\Fig~\ref{fig:lpt_comparison}), suggesting the possibility that \src{}, too, is a WD-MD polar system.
However, its period of ${\sim}1.16\,$hours would make it shorter than the minimum known orbital period of $1.3\,$hours for confirmed polars \citep{schwope2025polarcatcatalogpolarslowaccretion}.
This makes it an interesting testbed for the suggestion advanced by \citet{2025A&A...695L...8R} that ``short LPTs'' ($P \lesssim 78\,$mins) and ``long LPTs'' ($P \gtrsim 78\,$mins) represent different classes of systems (AM CVns and CVs, respectively).
\begin{figure*}{th}
    \centering
    \includegraphics[width=0.98\linewidth]{lpt_comparison.pdf}
    \caption{Duty cycle vs radio pulse period for various long period radio emitters \citep{deRuiter2025,2022Natur.601..526H,2023Natur.619..487H,2024NatAs...8.1159C,2005Natur.434...50H,deRuiter2025,Lee2025}. The duty cycles are derived from the reported pulse widths for LPTs, which are compared to the range of the fitted, unscattered pulse widths of \src{} at 10\% of the peak. AR\,Sco \citep{2016Natur.537..374M,2018A&A...611A..66S} and other sources without well-defined duty cycles or periods \citep[e.g.][and references therein]{2021ApJ...920...45W} are not included on this plot.}
    \label{fig:lpt_comparison}
\end{figure*}

@@

Another possibility is that radio emission is only observed when the MD produces dense enough particle outflows, in the form of winds, to activate the emission mechanism in the vicinity of the WD.
Such outflows can last for months [CITATION??], and would not be strictly periodic.
If intermittency is of this kind, then the observed activity windows will be similarly aperiodic, or quasiperiodic.
Again, long-term monitoring can help distinguish between these scenarios.

@@

We have shown that \src{} is in fact an LPT, as conjectured in \citetalias{2024MNRAS.535..909D}).
Its period of ${\sim}1.16\,$hours is slightly shorter than the minimum known orbital period of optically confirmed polars, marking it as an interesting case study in the context of \citet{2025A&A...695L...8R}'s suggestion that `short'-LPTs and `long'-LPTs may correspond to different classes of cataclysmic variables.

A timing solution has been presented (\Eqn~\ref{eqn:timing}) which includes a term to account for scattering at low frequencies ($\lesssim 300\,$MHz).
Without it, ToA predictions may be early by up to several tens of seconds, which is a sizeable fraction of the intrinsic pulse width.

The three higher-frequency pulses for which polarimetry was observable exhibited unique single-pulse polarisation behaviours.
Only the original ASKAP detection exhibited a PA curve that was amenable to being modelled with the rotating vector model; the others show either flat, or slightly concave up PA curves (\Fig~\ref{fig:RM}).

We conjecture that if \src{} is a WD pulsar, its intermittency may be a result of the spin-orbital resonance being near, but not exactly, a small-integer ratio, creating a aliased beating pattern on a time scale of months.
The same explanation may be applied to other LPTs with activity cycles that last on the order of months.
An alternative explanation is that the radio emission is drived by particle wind outflows from the main sequence companion.
Long-term monitoring of LPTs will reveal whether the intermittency is periodic, and thus help discriminate between these two scenarios.
